---
title: 'Work Practices and Perceptions from Women Core Developers in OSS Communities'
author: Edna Dias Canedo, Rodrigo Bonifácio, Márcio Okimoto, Alexander Serebrenik, Gustavo Pinto, and Eduardo Monteiro
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.path='figures/', dev=c('png', 'pdf'))
require(quanteda)
require(sqldf)
require(readtext)
require(wordcloud)
require(xtable)
require(plotrix)
require(gdata)
require(lattice)
require(Hmisc)
require(reshape2)
require(nparcomp)
setwd(".")
```

## Script assessing the second research question

### (01) Laod the datasets 

To answer RQ2, we use the following datasets. 

   * commits_female_categorized.csv: contains information about the contributions from the set WCD (women core developers)
   * commits_male_s1_categorized.csv: contains information about the contributions from the set MCD1 (men core developers) 
   * commits_male_s2_categorized.csv: contains information about the contributions from the set MCD2 (men core developers)
   * commits_male_s3_categorized.csv: contains information about the contributions from the set MCD3 (men core developers)

We first load these datasets, show some basic information, and perform some data cleanup. 

```{r load}
f_commits_ds <- read.csv("../datasets/commits_female_categorized.csv", head=T, sep=";")
colnames(f_commits_ds)
nrow(f_commits_ds)

f_commits_ds["commit_date"] <- as.character(as.Date(strptime((
   f_commits_ds$last_modified), "%a, %d %b %Y %H:%M:%S")))

f_commits_ds$gender <- "women"
f_commits_ds$genderGroup <- "WCD"

m_commits_s1_ds <- read.csv("../datasets/commits_male_s1_categorized.csv", head=T, sep=";")
colnames(m_commits_s1_ds)
nrow(m_commits_s1_ds)

m_commits_s1_ds["commit_date"] <- as.character(as.Date(strptime((
   m_commits_s1_ds$last_modified), "%a, %d %b %Y %H:%M:%S")))

m_commits_s1_ds$gender <- "men"
m_commits_s1_ds$genderGroup <- "MCD1"


m_commits_s2_ds <- read.csv("../datasets/commits_male_s2_categorized.csv", head=T, sep=";")
colnames(m_commits_s2_ds)
nrow(m_commits_s2_ds)

m_commits_s2_ds["commit_date"] <- as.character(as.Date(strptime((
   m_commits_s2_ds$last_modified), "%a, %d %b %Y %H:%M:%S")))

m_commits_s2_ds$gender <- "men"
m_commits_s2_ds$genderGroup <- "MCD2"


m_commits_s3_ds <- read.csv("../datasets/commits_male_s3_categorized.csv", head=T, sep=";")
colnames(m_commits_s3_ds)
nrow(m_commits_s3_ds)

m_commits_s3_ds["commit_date"] <- as.character(as.Date(strptime((
   m_commits_s3_ds$last_modified), "%a, %d %b %Y %H:%M:%S")))

m_commits_s3_ds$gender <- "men"
m_commits_s3_ds$genderGroup <- "MCD3"

commits_ds <- rbind(f_commits_ds, m_commits_s1_ds)
commits_ds <- rbind(commits_ds, m_commits_s2_ds)
commits_ds <- rbind(commits_ds, m_commits_s3_ds)

colnames(commits_ds)
nrow(commits_ds)
sqldf("select count (distinct(commit_date)) from commits_ds")
head(commits_ds[,c("commit_date", "last_modified")])


sqldf("select count(distinct login) from f_commits_ds")

sqldf("select count(distinct login) from m_commits_s1_ds")

sqldf("select count(distinct login) from m_commits_s2_ds")

sqldf("select count(distinct login) from m_commits_s3_ds")

head(sqldf("select  login, count(*) from m_commits_s3_ds 
       group by login order by 2 desc"))

# remove an outilier contributor 

commits_ds <- commits_ds[commits_ds$login != "michaelni",] 

nrow(commits_ds) 
sqldf("select genderGroup, count(*) 
       from commits_ds 
      group by genderGroup")
```


### (02) Summarizing our dataset 

Let's summarize our dataset by core developer, and also preserving his / her gender and gender group. 

```{r consolidate}
commits_by_user <- sqldf("select login, gender, genderGroup,
                                 count(distinct hash_commit) commits,
                                 max(commit_date) max_date, 
                                 min(commit_date) min_date, 
                                 count(distinct commit_date) frequency, 
                                 sum(num_files) files ,
                                 sum(num_files_changed) files_changed,
                                 sum(num_files_deleted) files_deleted,
                                 sum(num_files_added) files_added,
                                 sum(num_files_renamed) files_renamed,
                                 sum(num_lines_changed) lines_changed,
                                 sum(num_lines_deleted) lines_deleted,
                                 sum(num_lines_added) lines_added
                          from commits_ds
                          group by login, gender, genderGroup")
                          

colnames(commits_by_user)
nrow(commits_by_user)
```

### (03) Hypothesis testing about the number of commits. 

Now we can verify if there is any difference between these groups, in terms of total number of commits. 

   * First showing the boxplts 
   
```{r boxplot1,fig.width=5, fig.height = 3.5}
bwplot(commits_by_user$commits~commits_by_user$genderGroup, 
       ylab="Total number of contributions (commits)",    
       par.settings = list(
                           axis.text = list(cex = 0.75),
                           par.xlab.text = list(cex = 0.75),
                           par.ylab.text = list(cex = 0.75)
                          ))

```

   * Second performing a multicomparison test (using the Dunnett's method)
   
```{r dunnet-commits}
c<-nparcomp(commits~genderGroup, data=commits_by_user, asy.method = "logit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)

summary(c)
```

### (04) Hypothesis testing about the frequency of commits. 

Now we can verify if there is any difference between these groups, in terms of frequency of commits. 

   * First computing the metrics and showing the boxplts 

```{r bw-frequency,fig.width=5, fig.height = 3.5}
commits_by_user["date_diff"] = as.numeric(difftime(as.Date(commits_by_user$max_date), 
                                                   as.Date(commits_by_user$min_date), units="days"))

commits_by_user["frequency2"] = 100*commits_by_user$frequency / commits_by_user$date_diff 

cor.test(commits_by_user$frequency2, commits_by_user$commits, method="spearman")

bwplot(commits_by_user$frequency2~commits_by_user$genderGroup, ylab="Frequency of commits", 
       par.settings = list(
                           axis.text = list(cex = 0.75),
                           par.xlab.text = list(cex = 0.75),
                           par.ylab.text = list(cex = 0.75)
                          ))

```

   * Second computing the summary of the observations
```{r freq-summary}
summary(commits_by_user[commits_by_user$gender == 'women',]$frequency2)
summary(commits_by_user[commits_by_user$gender == 'men',]$frequency2)
```

   * Third performing a multicomparison test (using the Dunnett's method)

```{r dunnet-frequency}
c<-nparcomp(frequency2~genderGroup, data=commits_by_user, asy.method = "probit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)


summary(c)
```


### (05) Hypothesis testing about the size of the contributions 

#### 5.1 Considering lines of code

```{r bw-line-size, fig.width= 7,fig.height = 3.5} 
lines <- commits_by_user[,c("login", "gender", "genderGroup", 
                            "lines_added", "lines_changed", "lines_deleted")]

lines <- melt(lines, id.vars = c("login", "gender", "genderGroup"))

lines <- lines[order(lines$variable),]

lines$variable <- gsub("lines_added", "Lines Added", lines$variable)
lines$variable <- gsub("lines_changed", "Lines Changed", lines$variable)
lines$variable <- gsub("lines_deleted", "Lines Deleted", lines$variable)


bwplot(log(value)~genderGroup|variable, data=lines, 
       ylab="Lines of Code (log scale)", layout=c(3,1), 
       par.settings = list(
                           axis.text = list(cex = 0.55),
                           par.xlab.text = list(cex = 0.75),
                           par.ylab.text = list(cex = 0.75)
                          ), 
       par.strip.text = list(cex = 0.65))

la <- lines[lines$variable == "Lines Added", ]
lc <- lines[lines$variable == "Lines Changed", ]
ld <- lines[lines$variable == "Lines Deleted", ]


c<-nparcomp(value~genderGroup, data=la, asy.method = "probit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)

summary(c)


c<-nparcomp(value~genderGroup, data=lc, asy.method = "probit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)


summary(c)

c<-nparcomp(value~genderGroup, data=ld, asy.method = "probit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)


summary(c)
```


#### 5.2 Considering number of files

```{r bw-file-size, fig.width= 7,fig.height = 3.5} 
files <- commits_by_user[,c("login", "gender", "genderGroup", 
                            "files_added", "files_changed", "files_deleted")]

files <- melt(files, id.vars = c("login", "gender", "genderGroup"))

files <- files[order(files$variable),]

files$variable <- gsub("files_added", "Files Added", files$variable)
files$variable <- gsub("files_changed", "Files Changed", files$variable)
files$variable <- gsub("files_deleted", "Files Deleted", files$variable)


bwplot(log(value)~genderGroup|variable, data=files,ylab="Number of Files (Log Scale)",layout=c(3,1))


fa <- files[files$variable == "Files Added", ]
fc <- files[files$variable == "Files Changed", ]
fd <- files[files$variable == "Files Deleted", ]

c<-nparcomp(value~genderGroup, data=fa, asy.method = "probit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)


summary(c)

c<-nparcomp(value~genderGroup, data=fc, asy.method = "probit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)


summary(c)


c<-nparcomp(value~genderGroup, data=fd, asy.method = "probit", 
            type = "Dunnett", control = "WCD", alternative = "two.sided", info = FALSE)


summary(c)

```

### (06) Hypothesis testing about the type of the contributions 

   * First, lets investigate the size of the comments in the commits. 
   
```{r bw-comments, fig.width=8, fig.height = 3.5}
commits_ds["comment_size"] = nchar(as.character(commits_ds$message))

# boxplot(log(commits_ds$comment_size)~commits_ds$genderGroup)

wilcox.test(comment_size~gender, data = commits_ds)
```

   * Second, lets investigate the type of the contributions. 

```{r bw-contribution, fig.width=8, fig.height = 3.5}

contributions_ds <- sqldf("select login, gender, genderGroup, class, count(*) total
                           from commits_ds 
                           group by login, gender, genderGroup, class")


bwplot(log(total)~genderGroup|class, data=contributions_ds,  par.settings = list(
                               axis.text = list(cex = 0.55),
                               par.xlab.text = list(cex = 0.75),
                               par.ylab.text = list(cex = 0.75)
                          ),  par.strip.text = list(cex = 0.65), layout=c(5,1))

chisq.test(commits_ds$class, commits_ds$gender)
```

There is a huge difference in the number of commits. Perhaps we should first balance the dataset. 

```{r bw-balanced-contribution, fig.width=10, fig.height = 3.5}
m_commits_s1_ds <- m_commits_s1_ds[sample(nrow(m_commits_s1_ds), 22326, replace = FALSE, prob = NULL),]

m_commits_s2_ds <- m_commits_s2_ds[sample(nrow(m_commits_s2_ds), 22326, replace = FALSE, prob = NULL),]

m_commits_s3_ds <- m_commits_s3_ds[sample(nrow(m_commits_s3_ds), 22326, replace = FALSE, prob = NULL),]

nrow(f_commits_ds)
nrow(m_commits_s1_ds)
nrow(m_commits_s2_ds)
nrow(m_commits_s3_ds)


commits_ds <- rbind(f_commits_ds, m_commits_s1_ds)
commits_ds <- rbind(commits_ds, m_commits_s2_ds)
commits_ds <- rbind(commits_ds, m_commits_s3_ds)

contributions_ds <- sqldf("select login, gender, genderGroup, class, count(*) total
                           from commits_ds 
                           group by login, gender, genderGroup, class")


bwplot(log(total)~genderGroup|class, data=contributions_ds,  par.settings = list(
                               axis.text = list(cex = 0.55),
                               par.xlab.text = list(cex = 0.75),
                               par.ylab.text = list(cex = 0.75)
                          ),  par.strip.text = list(cex = 0.65), layout=c(5,1))

chisq.test(commits_ds$class, commits_ds$gender)
```
