---
title: 'Work Practices and Perceptions from Women Core Developers in OSS Communities'
author: Edna Dias Canedo, Rodrigo Bonifácio, Márcio Okimoto, Alexander Serebrenik, Gustavo Pinto, and Eduardo Monteiro
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.path='figures/', dev=c('png', 'pdf'))

require(viridis)
require(prettydoc)
require(quanteda)
require(sqldf)
require(readtext)
require(wordcloud)
require(xtable)
require(plotrix)
require(gdata)
require(lattice)
require(Hmisc)
require(tidyverse)
require(ggplot2)
require(reshape2)
require(plyr)
require(knitr)
setwd(".")
```

## Script assessing the closed questions of the survey

### (01) Laod the datasets 

To answer RQ1, we use one dataset. 

   * responses: contains the answers to the Likert scale questions of the survey


```{r}
ds <- read.csv("../datasets/survey-closed-questions.csv", head=T, sep=",", na.strings=c(" "))
ds <- ds[, c(1, 2,4,30,31,32,33,34,35,5,6,7,8,9,10,11)]

colnames(ds) <- c("Id", "Age", 
                  "Experience", 
                  "Frequencty of Interaction with Other Members",
                  "Happyness with the contributions",
                  "Acceptance of Contributions",
                  "Rejections Due to Gender Bias", 
                  "Frequency of Rejections Due to Gender Bias", 
                  "Relevance of Gender Diversity in OSS Communities",
                  "Undergraduate Student",
                  "Bachelor Degree",
                  "Master Student",
                  "Master Degree", 
                  "PhD Student",
                  "PhD Degree", 
                  "Other Degree")

```


### (02) Demographics

#### Age


```{r}
ds$Age <- factor(ds$Age)
kable(count(ds, 'Age'))
```

#### Education

```{r}
education <- ds[,c(1, 10,11,12,13,14,15)]
colnames(education)
education <- melt(education, rm.na= T, id.vars=c("Id"), variable.name="Option", value.name="Degree")

education[education==""]<-NA

degree <- sqldf("select Degree, count(*) Total from education where Degree is not null group by Degree order by 2 desc")
kable(degree)
```

#### Years contributing to OSS communities

```{r}
ds1 <- ds[, c(1,3)]
experience_options <- read.csv("../datasets/experience_options.csv", head=T, sep=",")

ds1 <- merge(ds1, experience_options)
summary <- sqldf("select Option, Experience, count(*) Total from ds1 group by Option, Experience order by 1")
kable(summary)
```

### (03) How important is gender diversity in OSS Communities?

```{r}
ds1 <- ds[, c(1,9)]
colnames(ds1) <- c("Id", "RelevanceOfDiversity")
summary <- sqldf("select RelevanceOfDiversity, count(*) Total from ds1 group by RelevanceOfDiversity order by 2 desc")
kable(summary)
barplot(summary$Total, ylim=c(0, 25), col = viridis(nrow(summary)))
legend("topright", legend=summary$RelevanceOfDiversity, fill = viridis(nrow(summary)))
```

### (04) How often do you feel that your contributions were not well received due to your gender?

```{r}
frequency_options <- read.csv("../datasets/frequency_options.csv", head=T, sep=",")
ds1 <- ds[, c(1,8)]
colnames(ds1)
colnames(ds1) <- c("Id", "FrequencyId")

ds1 <- merge(ds1, frequency_options)

summary <- sqldf("select FrequencyId, Frequency, count(*) Total from ds1 group by FrequencyId, Frequency order by 1")
kable(summary)
barplot(summary$Total, ylim=c(0, 25), col = viridis(nrow(summary)))
legend("topright", legend=summary$Frequency, fill = viridis(nrow(summary)))
```

### (05) Are you happy with your participation in OSS Communities?

```{r}
ds1 <- ds[, c(1,5)]
colnames(ds1) <- c("Id", "Answer")
ds1$Answer <- ifelse(ds1$Answer == 0, "False", "True")
summary <- sqldf("select Answer as 'Are you happy with your participation?', count(*) Total from ds1 group by Answer")
kable(summary)
```